---
title: "3_DIfferential Analysis"
author: "Christian Ayala"
date: "3/19/2021"
output: html_document
---

This Notebook is to perform the differential analysis of the normalized AUC from *Compound Discoverer*.

# 1. Importing Libraries

```{r libraries, message=FALSE}
library(tidyverse)
library(ggpubr)
library(ggsci)
library(ggrepel)
library(pheatmap)
library(RColorBrewer)
source('functions_cdis_diff.R')
```

# 2. Import data

Because differential analysis includes the calculations of means, the data to be used in this section is the normalized untransformed data

```{r set_path, message=FALSE}
# set path variables
project_dir <- getwd()
project_name <- 'Bog_lbl_unlbl_1.7int'
figures_dir <- file.path(project_dir, paste0(project_name, '_output_figures'))
tables_dir <- file.path(project_dir,  paste0(project_name, '_output_tables'))
norm_auc_table_file <- file.path(tables_dir, 'normalized_untransformed_auc_table.csv')

# Load compounds_table

norm.matrix <- read_csv(norm_auc_table_file)
norm.matrix <- column_to_rownames(norm.matrix, var = 'X1')

# Import metadata and fix names
metadata_file <- file.path(tables_dir, 'fixed_metadata.csv')
metadata <- read_csv(metadata_file)

```

# 3. Calculate ratios and log2fold-change

## 3.1 Get samples per each treatment

```{r}

# Get the average AUC per each of the treatments
## Get samples per treatment

T0.samples <- get_samples(metadata, Treatment = 'Time', value = 'T0')
T1.samples <- get_samples(metadata, Treatment = 'Time', value = 'T1')
T2.samples <- get_samples(metadata, Treatment = 'Time', value = 'T2')
T3.samples <- get_samples(metadata, Treatment = 'Time', value = 'T3')

```

## 3.2 Get table of differentially expressed features per each comparison

```{r}

T1_to_T0.diff_table <- get_diff_table(norm.matrix, control.sample_list = T0.samples, treatment.sample_list = T1.samples)
T2_to_T0.diff_table <- get_diff_table(norm.matrix, control.sample_list = T0.samples, treatment.sample_list = T2.samples)
T3_to_T0.diff_table <- get_diff_table(norm.matrix, control.sample_list = T0.samples, treatment.sample_list = T3.samples)

```

# 4. Plots of differentially expressed features

## 4.1 Number of differentially expressed features in each group

```{r}

num_diff_features <- tibble(Comparison = c('T1_to_T0', 'T2_to_T0', 'T3_to_T0'),
                            count = 0)

num_diff_features[1, 2] <- sum(T1_to_T0.diff_table$pval < 0.05, na.rm = TRUE)
num_diff_features[2, 2] <- sum(T2_to_T0.diff_table$pval < 0.05, na.rm = TRUE)
num_diff_features[3, 2] <- sum(T3_to_T0.diff_table$pval < 0.05, na.rm = TRUE)

num_diff_plot <- plot_col(num_diff_features, count, Comparison)
num_diff_plot

figure_file <- file.path(figures_dir, 'Num_diff_features.png')
ggsave(figure_file, num_diff_plot, dpi = 300)

```

## 4.2 Volcano plots

```{r}

lfc.t <- 1
pval.t <- 0.05

T1_to_T0_volcano <- plot_volcano(T1_to_T0.diff_table, log2FC, pval, lfc.t, pval.t) +
  labs(subtitle = 'T1 vs T0')
T1_to_T0_volcano

figure_file <- file.path(figures_dir, 'T1_T0_volcano.png')
ggsave(figure_file, T1_to_T0_volcano, dpi = 300)


T2_to_T0_volcano <- plot_volcano(T2_to_T0.diff_table, log2FC, pval, lfc.t, pval.t) +
  labs(subtitle = 'T2 vs T0')
T2_to_T0_volcano

figure_file <- file.path(figures_dir, 'T2_T0_volcano.png')
ggsave(figure_file, T2_to_T0_volcano, dpi = 300)

T3_to_T0_volcano <- plot_volcano(T3_to_T0.diff_table, log2FC, pval, lfc.t, pval.t) +
  labs(subtitle = 'T3 vs T0')
T3_to_T0_volcano

figure_file <- file.path(figures_dir, 'T3_T0_volcano.png')
ggsave(figure_file, T3_to_T0_volcano, dpi = 300)
```

## 4.3 Hierarchical clustering analysis using heatmaps

Normalized AUC for all features

```{r}

# Set sampleID as row.names to annotate heatmap

colData <- column_to_rownames(metadata, var = 'SampleID') 
mapcolor <- colorRampPalette(brewer.pal(11, 'RdYlBu'))(100)[100:1]

figure_file <- file.path(figures_dir, 'All_features_HTMP.png')

png(figure_file)
pheatmap(norm.matrix,
         clustering_distance_rows = 'correlation',
         clustering_distance_cols = 'correlation',
         scale = 'row',
         annotation_col = colData,
         color = mapcolor,
         show_rownames = FALSE,
         cutree_cols = 4,
         main = 'Heatmap normalized AUC (all detected features)'
         )
dev.off()

```

Most differentially expressed features

```{r}

# Extract significant features of each comparison

sig.features <- c(T1_to_T0.diff_table$FeatureID[T1_to_T0.diff_table$log2FC > 1 & T1_to_T0.diff_table$pval < 0.05],
                  T2_to_T0.diff_table$FeatureID[T2_to_T0.diff_table$log2FC > 1 & T2_to_T0.diff_table$pval < 0.05],
                  T3_to_T0.diff_table$FeatureID[T1_to_T0.diff_table$log2FC > 1 & T3_to_T0.diff_table$pval < 0.05])

sig.features <- unique(sig.features)

sig.features_table <- norm.matrix[row.names(norm.matrix) %in% sig.features, ]
sig.features_table <- as.matrix(sig.features_table)

figure_file <- file.path(figures_dir, 'DE_features_HTMP.png')

png(figure_file)
pheatmap(sig.features_table,
         clustering_distance_rows = 'correlation',
         clustering_distance_cols = 'correlation',
         scale = 'row',
         annotation_col = colData,
         color = mapcolor,
         cutree_cols = 4,
         main = 'Heatmap normalized AUC (Differentialy expressed)'
         )
dev.off()

```

Log2FC for all the comparisons

```{r}

# Extract log2FC of all the studied comparisons

log2fc.table <- data.frame(cbind(T1_to_T0.diff_table$log2FC, T2_to_T0.diff_table$log2FC, T3_to_T0.diff_table$log2FC))
colnames(log2fc.table) <- c('T1_to_T0', 'T2_to_T0', 'T3_to_T0')
rownames(log2fc.table) <- row.names(norm.matrix)

# Removing rows with NA and Inf values
log2fc.table <- drop_na(log2fc.table)
log2fc.table <- log2fc.table[is.finite(rowSums(log2fc.table)),]

log2fc.table <- as.matrix(log2fc.table)

figure_file <- file.path(figures_dir, 'log2FC_HTMP.png')
png(figure_file)
pheatmap(log2fc.table,
         clustering_distance_rows = 'correlation',
         clustering_distance_cols = 'correlation',
         show_rownames = FALSE,
         color = mapcolor,
         main = 'Heatmap log2FC (all features)'
         )
dev.off()

```


