---
title: "3_Differential Analysis"
author: "Christian Ayala"
date: "3/19/2021"
output: html_document
---

This Notebook is to perform the differential analysis of the normalized AUC from *Compound Discoverer*.

# 1. Importing Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggrepel)
library(ggsci)
library(pheatmap)
library(RColorBrewer)
library(readxl)
library(UpSetR)
source('functions_cdis_diff.R')
```

# 2. Import data

Set if the data to be used is going to be labeled or unlabeled

```{r}
# Flag for labeled / unlabeled data, set TRUE or FALSE

label = TRUE
```

Because differential analysis includes the calculations of means, the data to be used in this section is the normalized untransformed data

```{r set_path, message=FALSE, warning=FALSE}
# set path variables
project_dir <- getwd()
project_name <- 'Bog_1e5_label'
figures_dir <- file.path(project_dir, paste0(project_name, '_output_figures'))
tables_dir <- file.path(project_dir,  paste0(project_name, '_output_tables'))
norm_auc_table_file <- file.path(tables_dir, 'normalized_untransformed_auc_table.csv')

# If the flag was setted before, nothing needs to be changed here, the correct file will be automatically used
if(label == TRUE){
  compounds_table_file <- file.path(tables_dir, 'compounds_table.csv')
}else{
  compounds_table_file <- file.path(tables_dir, 'gap_filled_compounds_table.csv')
}

# Load auc_table

norm.matrix <- read_csv(norm_auc_table_file)
norm.matrix <- column_to_rownames(norm.matrix, var = 'X1')

# Load compounds table and add a column that has names for identified compounds and FeatureID for the rest (for plotting purpouses)

compounds_table <- read_csv(compounds_table_file)

# Import metadata and fix names
metadata_file <- file.path(tables_dir, 'fixed_metadata.csv')
metadata <- read_csv(metadata_file)

# Import NMR data
nmr_file <- file.path(project_dir, 'data', 'Saleska_BOG_concentrations_ALL.xlsx')
nmr_table <- read_xlsx(nmr_file, skip = 4)
colnames(nmr_table) <- str_remove(colnames(nmr_table), '_1D.*')
colnames(nmr_table)[1] <- 'Name'

```

# 3. Calculate ratios and log2fold-change

## 3.1 Get samples per each treatment

```{r}

# Get the average AUC per each of the treatments
## Get samples per treatment

T0.samples <- get_samples(metadata, Treatment = 'Time', value = 'T0')
T1.samples <- get_samples(metadata, Treatment = 'Time', value = 'T1')
T2.samples <- get_samples(metadata, Treatment = 'Time', value = 'T2')
T3.samples <- get_samples(metadata, Treatment = 'Time', value = 'T3')

```

## 3.2 Get table of differentially expressed features per each comparison

```{r}

# The following function will calculate ratio, log2FC, p values and adjusted pvalues. If no replicates are available for EACH treatment
# please use the get_diff_table_no_pval() function

T1_to_T0.diff_table <- get_diff_table(norm.matrix, control.sample_list = T0.samples, treatment.sample_list = T1.samples)
T2_to_T0.diff_table <- get_diff_table(norm.matrix, control.sample_list = T0.samples, treatment.sample_list = T2.samples)
T3_to_T0.diff_table <- get_diff_table(norm.matrix, control.sample_list = T0.samples, treatment.sample_list = T3.samples)
```

Merge differentially expressed features with the annotation

```{r}

# Create dataframes for the up and downregulated metabolites at each time point and merge them with the compound information

T1_to_T0.diff_table <- compounds_table %>% 
  select(-contains('Annotation Source:'), - contains('Results'), -SampleID, -AUC, -Label, -Material, - Time) %>% 
  right_join(T1_to_T0.diff_table, by = 'FeatureID') %>% 
  distinct() %>% 
  filter(!is.nan(log2FC)) %>% 
  mutate(Comment = case_when(log2FC == Inf ~ 'Not present in control',
                             log2FC == -Inf ~ 'Only present in control',
                             log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

T1_to_T0.diff_table$Comment <- factor(T1_to_T0.diff_table$Comment, 
                               levels = c('Only present in control', 'Downregulated', 'Upregulated', 'Not present in control' ))

table_file <- file.path(tables_dir, 'Diff_expressed_T1.csv')
write_csv(T1_to_T0.diff_table, table_file )

T2_to_T0.diff_table <- compounds_table %>% 
  select(-contains('Annotation Source:'), - contains('Results'), -SampleID, -AUC, -Label, -Material, - Time) %>%
  right_join(T2_to_T0.diff_table, by = 'FeatureID') %>% 
  distinct() %>% 
  filter(!is.nan(log2FC)) %>% 
  mutate(Comment = case_when(log2FC == Inf ~ 'Not present in control',
                             log2FC == -Inf ~ 'Only present in control',
                             log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

T2_to_T0.diff_table$Comment <- factor(T2_to_T0.diff_table$Comment, 
                               levels = c('Only present in control', 'Downregulated', 'Upregulated', 'Not present in control' ))

table_file <- file.path(tables_dir, 'Diff_expressed_T2.csv')
write_csv(T2_to_T0.diff_table, table_file )

T3_to_T0.diff_table <- compounds_table %>% 
  select(-contains('Annotation Source:'), - contains('Results'), -SampleID, -AUC, -Label, -Material, - Time) %>% 
  right_join(T3_to_T0.diff_table, by = 'FeatureID') %>% 
  distinct() %>% 
  filter(!is.nan(log2FC)) %>% 
  mutate(Comment = case_when(log2FC == Inf ~ 'Not present in control',
                             log2FC == -Inf ~ 'Only present in control',
                             log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

T3_to_T0.diff_table$Comment <- factor(T3_to_T0.diff_table$Comment, 
                               levels = c('Only present in control', 'Downregulated', 'Upregulated', 'Not present in control' ))

table_file <- file.path(tables_dir, 'Diff_expressed_T3.csv')
write_csv(T3_to_T0.diff_table, table_file )

```

Extract most differentially expressed features

```{r warning=FALSE}

# Extract significant features of each comparison based on adjusted pvalue
sig_features <- c(T1_to_T0.diff_table$FeatureID[T1_to_T0.diff_table$pval.adj < 0.05],
                  T2_to_T0.diff_table$FeatureID[T2_to_T0.diff_table$pval.adj < 0.05],
                  T3_to_T0.diff_table$FeatureID[T3_to_T0.diff_table$pval.adj < 0.05])

sig_features <- unique(sig_features)
```


# 4. Plots of dysregulated features

## 4.1 Number of dysregulated features in each group

```{r}

num_diff_features <- tibble(Comparison = rep(c('T1_to_T0', 'T2_to_T0', 'T3_to_T0'), each = 2),
                            Type = rep(c('Name', 'No name'), 3),
                            count = 0)
num_diff_features$Type <- factor(num_diff_features$Type, levels = c('No name', 'Name'))

num_diff_features[1, 3] <- sum(T1_to_T0.diff_table$pval.adj < 0.05 & !is.na(T1_to_T0.diff_table$Name), na.rm = TRUE)
num_diff_features[2, 3] <- sum(T1_to_T0.diff_table$pval.adj < 0.05 & is.na(T1_to_T0.diff_table$Name), na.rm = TRUE)
num_diff_features[3, 3] <- sum(T2_to_T0.diff_table$pval.adj < 0.05 & !is.na(T2_to_T0.diff_table$Name), na.rm = TRUE)
num_diff_features[4, 3] <- sum(T2_to_T0.diff_table$pval.adj < 0.05 & is.na(T2_to_T0.diff_table$Name), na.rm = TRUE)
num_diff_features[5, 3] <- sum(T3_to_T0.diff_table$pval.adj < 0.05 & !is.na(T3_to_T0.diff_table$Name), na.rm = TRUE)
num_diff_features[6, 3] <- sum(T3_to_T0.diff_table$pval.adj < 0.05 & is.na(T3_to_T0.diff_table$Name), na.rm = TRUE)


num_diff_plot <- plot_col(num_diff_features, count, Comparison, Type)
num_diff_plot

figure_file <- file.path(figures_dir, 'Num_diff_features.png')
ggsave(figure_file, num_diff_plot, dpi = 300)
```

## 4.2 Dysregulated features shared at different timepoints

An **upset plot** is used instead of a Venn Diagram asi it provides a better visualization of the number of figures that are being shared among groups

```{r}
# Filter diff tables by pval.adj

T1.sig <- T1_to_T0.diff_table %>% 
  filter(pval.adj < 0.05)
T2.sig <- T2_to_T0.diff_table %>% 
  filter(pval.adj < 0.05)
T3.sig <- T3_to_T0.diff_table %>% 
  filter(pval.adj < 0.05)

# Get list of features at each time point for each of the cases
T1.upregulated <- get_vectors(T1.sig, 'Comment', 'Upregulated', 'FeatureID')
T1.downregulated <- get_vectors(T1.sig, 'Comment', 'Downregulated', 'FeatureID')
T1.not_in_control <- get_vectors(T1.sig, 'Comment', 'Not present in control', 'FeatureID')
T1.only_in_control <- get_vectors(T1.sig, 'Comment', 'Only present in control', 'FeatureID')

T2.upregulated <- get_vectors(T2.sig, 'Comment', 'Upregulated', 'FeatureID')
T2.downregulated <- get_vectors(T2.sig, 'Comment', 'Downregulated', 'FeatureID')
T2.not_in_control <- get_vectors(T2.sig, 'Comment', 'Not present in control', 'FeatureID')
T2.only_in_control <- get_vectors(T2.sig, 'Comment', 'Only present in control', 'FeatureID')

T3.upregulated <- get_vectors(T3.sig, 'Comment', 'Upregulated', 'FeatureID')
T3.downregulated <- get_vectors(T3.sig, 'Comment', 'Downregulated', 'FeatureID')
T3.not_in_control <- get_vectors(T3.sig, 'Comment', 'Not present in control', 'FeatureID')
T3.only_in_control <- get_vectors(T3.sig, 'Comment', 'Only present in control', 'FeatureID')
```

```{r fig.height=8}
upset_input <- list(T1.upregulated = T1.upregulated,
                    T1.downregulated = T1.downregulated,
                    T1.not_in_control = T1.not_in_control,
                    T1.only_in_control = T1.only_in_control,
                    T2.upregulated = T2.upregulated,
                    T2.downregulated = T2.downregulated,
                    T2.not_in_control = T2.not_in_control,
                    T2.only_in_control = T2.only_in_control,
                    T3.upregulated = T3.upregulated,
                    T3.downregulated = T3.downregulated,
                    T3.not_in_control = T3.not_in_control,
                    T3.only_in_control = T3.only_in_control)

upset_metadata <- data.frame(sets = as.vector(names(upset_input)),
                         Time = rep(c('T1', 'T2', 'T3'), each = 4))

upset_features <- upset(fromList(upset_input), order.by = "freq", cutoff = 1, nsets = 12,
                    mainbar.y.label = 'Number of shared features', sets.x.label = 'Number of features',
                    text.scale = c(1.5, 1, 1.5, 1, 1.3, 1.3),
                    set.metadata = list(data = upset_metadata, 
                                        plots = list(list(type = 'matrix_rows', column = 'Time', 
                                                          colors = c(T1 = '#EFC000FF', T2 = '#868686FF', T3 = '#CD534CFF')))))
upset_features

figure_file <- file.path(figures_dir, 'Upset-KEGG.png')
png(figure_file, width = 800, height = 800)
upset_features
dev.off()
```

## 4.2 Volcano plots

```{r warning=FALSE}

lfc.t <- 2
pval.t <- 0.05

T1_to_T0_volcano <- plot_volcano(T1_to_T0.diff_table, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'T1 vs T0')
T1_to_T0_volcano

figure_file <- file.path(figures_dir, 'T1_T0_volcano.png')
ggsave(figure_file, T1_to_T0_volcano, dpi = 300)


T2_to_T0_volcano <- plot_volcano(T2_to_T0.diff_table, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'T2 vs T0')
T2_to_T0_volcano

figure_file <- file.path(figures_dir, 'T2_T0_volcano.png')
ggsave(figure_file, T2_to_T0_volcano, dpi = 300)

T3_to_T0_volcano <- plot_volcano(T3_to_T0.diff_table, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'T3 vs T0')
T3_to_T0_volcano

figure_file <- file.path(figures_dir, 'T3_T0_volcano.png')
ggsave(figure_file, T3_to_T0_volcano, dpi = 300)
```

## 4.3 Van Krevelen Diagrams of dysregulated features

**Dysregulated features at T1**

```{r}

T1_vank <- plot_vank(T1.sig, Class, Comment)
T1_vank

figure_file <- file.path(figures_dir, 'DE_vank_T1.png')
ggsave(figure_file, T1_vank)
```

**Dysregulated features at T2**

```{r}
T2_vank <- plot_vank(T2.sig, Class, Comment)
T2_vank

figure_file <- file.path(figures_dir, 'DE_vank_T2.png')
ggsave(figure_file, T2_vank)
```

**Dysregulated features at T3**

```{r}
T3_vank <- plot_vank(T3.sig, Class, Comment)
T3_vank

figure_file <- file.path(figures_dir, 'DE_vank_T3.png')
ggsave(figure_file, T3_vank)
```

## 4.4 GFE plots of dysregulated features

**Dysregulated features at T1**

```{r}

my_colors = c('Not present in control' = '#d8365e', 
              'Upregulated' = '#d34849',
              'Downregulated' = '#005193',
              'Only present in control' = '#4b70de')

T1_GFE_box <- plot_boxplot(T1.sig, Comment, GFE, Comment, my_colors)
T1_GFE_box

figure_file <- file.path(figures_dir, 'DE_GFE_T1.png')
ggsave(figure_file, T1_GFE_box)
```

**Dysregulated features at T2**

```{r}
T2_GFE_box <- plot_boxplot(T2.sig, Comment, GFE, Comment, my_colors)
T2_GFE_box

figure_file <- file.path(figures_dir, 'DE_GFE_T2.png')
ggsave(figure_file, T2_GFE_box)
```

**Dysregulated features at T3**

```{r}
T3_GFE_box <- plot_boxplot(T3.sig, Comment, GFE, Comment, my_colors)
T3_GFE_box

figure_file <- file.path(figures_dir, 'DE_GFE_T3.png')
ggsave(figure_file, T3_GFE_box)
```


# 5. Hierarchical clustering analysis using heatmaps

Multiple heatmaps will be plotted, based on the differential expression and if the compounds were or not assigned and structure

### Normalized Area under the curve (AUC) heatmap for all detected features

```{r warning=FALSE}

# Initialize graphical device
dev.off()

# Set sampleID as row.names to annotate heatmap

col_annot <- column_to_rownames(metadata, var = 'SampleID') 
mapcolor <- colorRampPalette(brewer.pal(11, 'RdYlBu'))(100)[100:1]

figure_file <- file.path(figures_dir, 'All_features_heatmap.pdf')

annot_colors <- list(
  Time = c(T0 ='#0073C2FF', T1 = '#EFC000FF', T2 = '#868686FF', T3 = '#CD534CFF'),
  Material = c(Litter = 'green4', Litter_Peat = 'chocolate4'),
  Origin = c(LC_MS2 = '#0086cb', NMR = '#d23936'))

pdf(figure_file)
pheatmap(norm.matrix,
         clustering_distance_rows = 'correlation',
         clustering_distance_cols = 'correlation',
         scale = 'row',
         annotation_col = col_annot,
         annotation_colors = annot_colors,
         color = mapcolor,
         show_rownames = FALSE,
         cutree_cols = 5,
         main = 'All features (scaled AUC)'
)
dev.off()
```

[Scaled Area under the curve (AUC) heatmap for all detected features](`r toString(figure_file)`)

### Normalized Area under the curve (AUC) heatmap for only features that were identified

```{r}
# Extract all features that have names
named_compounds <- compounds_table %>% 
  select(FeatureID, Name, name4plot) %>% 
  filter(!is.na(Name)) %>% 
  distinct()

# Create a matrix of only identified features

named_compounds.matrix <- norm.matrix[rownames(norm.matrix) %in% named_compounds$FeatureID,]
row.names(named_compounds.matrix) <- compounds_table$name4plot[match(row.names(named_compounds.matrix), compounds_table$FeatureID)]

```

```{r warning=FALSE}

figure_file <- file.path(figures_dir, 'All_identified_features_heatmap.pdf')
dev.off()
pdf(figure_file, width = 15, height = 10)
pheatmap(named_compounds.matrix,
         clustering_distance_rows = 'correlation',
         clustering_distance_cols = 'correlation',
         scale = 'row',
         annotation_col = col_annot,
         annotation_colors = annot_colors,
         color = mapcolor,
         cutree_cols = 4,
         cutree_rows = 5,
         fontsize_row = 6,
         main = 'Identified features (scaled AUC)'
)
dev.off()
```

[Scaled Area under the curve (AUC) heatmap for only features that were identified](`r toString(figure_file)`)

### All dysregulated features

```{r}
# Create an AUC matrix that contains only features that are differentially expressed

sig_features.matrix <- norm.matrix[row.names(norm.matrix) %in% sig_features, ]

row.names(sig_features.matrix) <- compounds_table$name4plot[match(row.names(sig_features.matrix), compounds_table$FeatureID)]
```

```{r }
figure_file <- file.path(figures_dir, 'Dysreg_features_heatmap.pdf')

dev.off()
pdf(figure_file, width = 10, height = 10)
pheatmap(sig_features.matrix,
         clustering_distance_rows = 'correlation',
         clustering_distance_cols = 'correlation',
         scale = 'row',
         annotation_col = col_annot,
         annotation_colors = annot_colors,
         color = mapcolor,
         cutree_cols = 4,
         fontsize_row = 6,
         main = 'Dysregulated features (scaled AUC)'
)
dev.off()
```

[Scaled Area under the curve (AUC) heatmap for all dysregulated features](`r toString(figure_file)`)

### Dysregulated features that were identified

```{r }
named_sig_features.matrix <- sig_features.matrix[rownames(sig_features.matrix) %in% named_compounds$Name, ]

figure_file <- file.path(figures_dir, 'Dysreg_identified_features.pdf')

dev.off()
pdf(figure_file, width = 10)
pheatmap(named_sig_features.matrix,
         clustering_distance_rows = 'correlation',
         clustering_distance_cols = 'correlation',
         scale = 'row',
         annotation_col = col_annot,
         annotation_colors = annot_colors,
         color = mapcolor,
         cutree_cols = 4,
         fontsize_row = 8,
         main = 'Identified dysregulated features (scaled AUC)'
)
dev.off()
```

[Scaled Area under the curve (AUC) heatmap for identified dysregulated features](`r toString(figure_file)`)

# 6. Hierarchical clustering analysis with added NMR data

## 6.1 Differential analysis of NMR data

```{r}

# Create a matrix with NMR data
nmr.matrix <- nmr_table %>% 
  column_to_rownames(var = 'Name') %>% 
  select(metadata$SampleID)

# The following function will calculate ratio, log2FC, p values and adjusted pvalues. If no replicates are available for EACH treatment
# please use the get_diff_table_no_pval() function

T1_to_T0.nmr <- get_diff_table(nmr.matrix, control.sample_list = T0.samples, treatment.sample_list = T1.samples)
T2_to_T0.nmr <- get_diff_table(nmr.matrix, control.sample_list = T0.samples, treatment.sample_list = T2.samples)
T3_to_T0.nmr <- get_diff_table(nmr.matrix, control.sample_list = T0.samples, treatment.sample_list = T3.samples)

```


```{r}
# Join differential NMR data with NMR table
T1_to_T0.nmr <- nmr_table %>%
  select(Name, `KEGG Compound ID`, Formula, Weight, InChI) %>% 
  right_join(T1_to_T0.nmr, by = c('Name' = 'FeatureID')) %>% 
  filter(!is.nan(log2FC)) %>% 
  mutate(Comment = case_when(log2FC == Inf ~ 'Not present in control',
                             log2FC == -Inf ~ 'Only present in control',
                             log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

T1_to_T0.nmr$Comment <- factor(T1_to_T0.nmr$Comment, 
                               levels = c('Only present in control', 'Downregulated', 'Upregulated', 'Not present in control' ))

table_file <- file.path(tables_dir, 'Diff_expressed_NMR_T1.csv')
write_csv(T1_to_T0.nmr, table_file )

T2_to_T0.nmr <- nmr_table %>%
  select(Name, `KEGG Compound ID`, Formula, Weight, InChI) %>% 
  right_join(T2_to_T0.nmr, by = c('Name' = 'FeatureID')) %>% 
  filter(!is.nan(log2FC)) %>% 
  mutate(Comment = case_when(log2FC == Inf ~ 'Not present in control',
                             log2FC == -Inf ~ 'Only present in control',
                             log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

T2_to_T0.nmr$Comment <- factor(T2_to_T0.nmr$Comment, 
                               levels = c('Only present in control', 'Downregulated', 'Upregulated', 'Not present in control' ))

table_file <- file.path(tables_dir, 'Diff_expressed_NMR_T2.csv')
write_csv(T2_to_T0.nmr, table_file )

T3_to_T0.nmr <- nmr_table %>%
  select(Name, `KEGG Compound ID`, Formula, Weight, InChI) %>% 
  right_join(T3_to_T0.nmr, by = c('Name' = 'FeatureID')) %>% 
  filter(!is.nan(log2FC)) %>% 
  mutate(Comment = case_when(log2FC == Inf ~ 'Not present in control',
                             log2FC == -Inf ~ 'Only present in control',
                             log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

T3_to_T0.nmr$Comment <- factor(T3_to_T0.nmr$Comment, 
                               levels = c('Only present in control', 'Downregulated', 'Upregulated', 'Not present in control' ))

table_file <- file.path(tables_dir, 'Diff_expressed_NMR_T3.csv')
write_csv(T3_to_T0.nmr, table_file )
```

## 6.2 Heatmaps including NMR data

```{r}
# Join NMR data with dysregulated LC data
sig_nmr.matrix <- rbind(sig_features.matrix, nmr.matrix)
```

```{r}

row_annot <- tibble(Name = row.names(sig_nmr.matrix)) %>% 
  mutate(Origin = ifelse(Name %in% row.names(nmr.matrix), 'NMR', 'LC_MS2')) %>% 
  column_to_rownames(var = 'Name')

figure_file <- file.path(figures_dir, 'Dysreg_features_and_NMR.pdf')
dev.off()
pdf(figure_file, width = 15, height = 15)
pheatmap(sig_nmr.matrix,
         clustering_distance_rows = 'correlation',
         clustering_distance_cols = 'correlation',
         scale = 'row',
         annotation_col = col_annot,
         annotation_row = row_annot,
         annotation_colors = annot_colors,
         color = mapcolor,
         cutree_cols = 4,
         cutree_rows = 5,
         fontsize_row = 6,
         main = 'Dysregulated features and NMR data'
)
dev.off()
```

[Dysregulated features and NMR data](`r toString(figure_file)`)

```{r}
# Join NMR data with dysregulated LC data
named_sig_nmr.matrix <- rbind(named_sig_features.matrix, nmr.matrix)
```

```{r}
figure_file <- file.path(figures_dir, 'Dysreg_identified_features_and_NMR.pdf')
dev.off()
pdf(figure_file, width = 15, height = 10)
pheatmap(named_sig_nmr.matrix,
         clustering_distance_rows = 'correlation',
         clustering_distance_cols = 'correlation',
         scale = 'row',
         annotation_col = col_annot,
         annotation_row = row_annot,
         annotation_colors = annot_colors,
         color = mapcolor,
         cutree_cols = 4,
         cutree_rows = 5,
         fontsize_row = 6,
         main = 'Identified dysregulated features and NMR data'
)
dev.off()
```

[Identified dysregulated features and NMR data](`r toString(figure_file)`)
