---
title: "4_Dysregulated Pathways"
author: "Christian Ayala"
date: "4/14/2021"
output: html_document
---

This Notebook will explore the KEGG (and Metabolika) Pathways that are dysregulated

# 1. Importing Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(ggpubr)
library(ggsci)
library(ggrepel)
library(pheatmap)
library(RColorBrewer)
library(scales)
library(plyr)
library(cowplot)
library(readxl)
library(KEGGREST)
library(tidyverse)
source('functions_dysreg_path.R')
```

# 2. Import Data

```{r}
# Flag for labeled / unlabeled data, set TRUE or FALSE

label = TRUE
```

```{r set_path, message=FALSE, warning=FALSE}
# set path variables
project_dir <- getwd()
project_name <- 'Bog_1e5_label'
figures_dir <- file.path(project_dir, paste0(project_name, '_output_figures'))
tables_dir <- file.path(project_dir,  paste0(project_name, '_output_tables'))
```

This analysis uses the tables generated during the Differential Analysis script.

```{r message=FALSE, warning=FALSE}
# Load compounds table

# If the flag was set before, nothing needs to be changed here, the correct file will be automatically used
if(label == TRUE){
  compounds_table_file <- file.path(tables_dir, 'compounds_table.csv')
}else{
  compounds_table_file <- file.path(tables_dir, 'gap_filled_compounds_table.csv')
}
compounds_table <- read_csv(compounds_table_file)

# Import metadata and fix names
metadata_file <- file.path(tables_dir, 'fixed_metadata.csv')
metadata <- read_csv(metadata_file)

# Import table(s) obtained during the Differential analysis
table_file <- file.path(tables_dir, 'Diff_expressed_T1.csv')
T1.diff_table <- read_csv(table_file)
table_file <- file.path(tables_dir, 'Diff_expressed_T2.csv')
T2.diff_table <- read_csv(table_file)
table_file <- file.path(tables_dir, 'Diff_expressed_T3.csv')
T3.diff_table <- read_csv(table_file)

# Import diff tables of NMR data
# Import table(s) obtained during the Differential analysis
table_file <- file.path(tables_dir, 'Diff_expressed_NMR_T1.csv')
T1.NMR.diff_table <- read_csv(table_file)
table_file <- file.path(tables_dir, 'Diff_expressed_NMR_T2.csv')
T2.NMR.diff_table <- read_csv(table_file)
table_file <- file.path(tables_dir, 'Diff_expressed_NMR_T3.csv')
T3.NMR.diff_table <- read_csv(table_file)

```

From Compound Discover export a **2-level** table with: 

- *KEGG Pathways* in the first level with the columns: PathwayID, Pathway Name
- *Compounds* in the second level (Only way I found to be able to export the **KEGG Compounds IDs**) with the columns: Checked, Molecular Weight, Name, Formula, KEGG Compound Ids, KEGG Compound Names

```{r message=FALSE, warning=FALSE}
kegg_table_file <- file.path(tables_dir, 'KEGG Pathways_ok.xlsx')
kegg_table <- read_xlsx(kegg_table_file)

# Fix column names
colnames(kegg_table) <- c('PathwayID', 'Pathway_Name', 'Name', 'Formula', 'Molecular Weight', 'KEGG_ID', 'KEGG_Name')

# Fix 2-level table into regular table

kegg_table <- kegg_table %>% 
  filter(Pathway_Name != 'Checked') %>% 
  mutate(Pathway_Name = ifelse(Pathway_Name == 'FALSE', NA, Pathway_Name)) %>% 
  fill(PathwayID, Pathway_Name, .direction = 'down') %>% 
  filter(!is.na(`Molecular Weight`)) %>% 
  select(-`Molecular Weight`)
```

# 3. Analyze KEGG Pathways annotated at each time point

Merge KEGG annotation with Differential analysis data

```{r}
# This step includes filtering for significant DE metabolites (pval.adj < 0.05)
T1.kegg <- T1.diff_table %>%
  select(FeatureID, Name, Formula, log2FC, pval.adj, Comment) %>% 
  left_join(kegg_table, by = c('Name', 'Formula')) %>% 
  distinct() %>% 
  filter(pval.adj < 0.05) %>% 
  filter(!is.na(Pathway_Name)) %>% 
  mutate(Time = 'T1')

T2.kegg <- T2.diff_table %>%
  select(FeatureID, Name, Formula, log2FC, pval.adj, Comment) %>% 
  left_join(kegg_table, by = c('Name', 'Formula')) %>% 
  distinct() %>% 
  filter(pval.adj < 0.05)%>% 
  filter(!is.na(Pathway_Name)) %>% 
  mutate(Time = 'T2')

T3.kegg <- T3.diff_table %>%
  select(FeatureID, Name, Formula, log2FC, pval.adj, Comment) %>% 
  left_join(kegg_table, by = c('Name', 'Formula')) %>% 
  distinct() %>% 
  filter(pval.adj < 0.05)%>% 
  filter(!is.na(Pathway_Name)) %>% 
  mutate(Time = 'T3')
```

## 3.1 Plot number of KEGG pathways

```{r}
# Set colors for plot
my_colors = c('Not present in control' = '#d8365e', 
              'Upregulated' = '#d34849',
              'Downregulated' = '#005193',
              'Only present in control' = '#4b70de')

T1_counts.kegg <- T1.kegg %>% 
  group_by(Comment) %>% 
  count(Pathway_Name) %>% 
  ungroup() %>% 
  mutate(Time = 'T1')

T1_counts.kegg.plot <- plot_col(T1_counts.kegg, n, Pathway_Name, Comment, my_colors = my_colors) +
  facet_wrap(~Comment) +
  labs(title = 'Dysregulated KEGG Pathways at T1',
       y = 'Pathway Name') +
  scale_x_continuous(breaks = pretty_breaks(3))
T1_counts.kegg.plot

figure_file <- file.path(figures_dir, 'Dysreg_KEGG_T1.png')
ggsave(figure_file, T1_counts.kegg.plot, dpi = 300)

T2_counts.kegg <- T2.kegg %>% 
  group_by(Comment) %>% 
  count(Pathway_Name) %>% 
  ungroup() %>% 
  mutate(Time = 'T2')

T2_counts.kegg.plot <- plot_col(T2_counts.kegg, n, Pathway_Name, Comment, my_colors = my_colors) +
  facet_wrap(~Comment) +
  labs(title = 'Dysregulated KEGG Pathways at T2',
       y = 'Pathway Name') +
  scale_x_continuous(breaks = pretty_breaks(1))
T2_counts.kegg.plot

figure_file <- file.path(figures_dir, 'Dysreg_KEGG_T2.png')
ggsave(figure_file, T2_counts.kegg.plot, dpi = 300)

T3_counts.kegg <- T3.kegg %>% 
  group_by(Comment) %>% 
  count(Pathway_Name) %>% 
  ungroup() %>% 
  mutate(Time = 'T3')

T3_counts.kegg.plot <- plot_col(T3_counts.kegg, n, Pathway_Name, Comment, my_colors = my_colors) +
  facet_wrap(~Comment) +
  labs(title = 'Dysregulated KEGG Pathways at T3',
       y = 'Pathway Name') +
  scale_x_continuous(breaks = pretty_breaks(2))
T3_counts.kegg.plot

figure_file <- file.path(figures_dir, 'Dysreg_KEGG_T3.png')
ggsave(figure_file, T3_counts.kegg.plot, dpi = 300)

```

## 3.2 General visualization of dysregulated KEGG pathways

```{r fig.height=10}

lcms_kegg <- rbind(T1_counts.kegg, T2_counts.kegg, T3_counts.kegg) %>% 
  mutate(Source = paste(Time, Comment, sep = '_')) %>% 
  select(-Comment, -Time)

lcms_kegg <- pivot_wider(lcms_kegg, names_from = Source, values_from = n)

lcms_kegg[is.na(lcms_kegg)] <- 0

lcms_kegg <- pivot_longer(lcms_kegg, cols = 2:6, names_to = 'Source', values_to = 'n')

lcms_kegg <- ddply(lcms_kegg, .(Source), transform, rescale = rescale(n))

kegg_heatmap <- ggplot(lcms_kegg,
                       aes(x = Source,
                           y = Pathway_Name,
                           fill = rescale)) +
  geom_tile(color = 'white') +
  coord_fixed() +
  scale_fill_distiller(palette = 'Blues', direction = 1) +
  theme_bw() +
  theme(plot.title = element_text(face = 'bold', hjust = 0.5),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        axis.title.x = element_blank()) +
  labs(title = 'Dysregulated KEGG Pathways',
       y = 'Pathway Name',
       fill = 'Scaled counts')
kegg_heatmap

figure_file <- file.path(figures_dir, 'KEGG_tile_heatmap.png')
ggsave(figure_file, kegg_heatmap, height = 10)
```

## 3.3 KEGG Pathways from NMR data

Use the KEGGREST API to find the *KEGG Pathways* associated with the *KEGG Compound ID* from each metabolite identified with NMR

```{r message=FALSE, warning=FALSE, results='hide'}

T1.NMR.diff_table <- filter(T1.NMR.diff_table, !is.na(`KEGG Compound ID`))

T1_NMR.kegg <- tibble(Pathway_Name = NA, KEGG_ID = NA, Name = NA, Comment = NA, Time = 'T1')
for(i in 1:length(T1.NMR.diff_table$`KEGG Compound ID`)){
  k_id <- T1.NMR.diff_table$`KEGG Compound ID`[i]
  query <- keggGet(k_id)
  pathway <- query[[1]]$PATHWAY
  if(is.null(pathway)){
    pathway <- NA
  }
  t_tibble <- tibble(Pathway_Name = pathway,
                     KEGG_ID = k_id,
                     Name = T1.NMR.diff_table$Name[i],
                     Comment = T1.NMR.diff_table$Comment[i],
                     Time = 'T1')
  T1_NMR.kegg <- rbind(T1_NMR.kegg, t_tibble)
  print(i)
}

T2.NMR.diff_table <- filter(T2.NMR.diff_table, !is.na(`KEGG Compound ID`))

T2_NMR.kegg <- tibble(Pathway_Name = NA, KEGG_ID = NA, Name = NA, Comment = NA, Time = 'T2')
for(i in 1:length(T2.NMR.diff_table$`KEGG Compound ID`)){
  k_id <- T2.NMR.diff_table$`KEGG Compound ID`[i]
  query <- keggGet(k_id)
  pathway <- query[[1]]$PATHWAY
  if(is.null(pathway)){
    pathway <- NA
  }
  t_tibble <- tibble(Pathway_Name = pathway,
                     KEGG_ID = k_id,
                     Name = T2.NMR.diff_table$Name[i],
                     Comment = T2.NMR.diff_table$Comment[i],
                     Time = 'T2')
  T2_NMR.kegg <- rbind(T2_NMR.kegg, t_tibble)
  print(i)
}

T3.NMR.diff_table <- filter(T3.NMR.diff_table, !is.na(`KEGG Compound ID`))

T3_NMR.kegg <- tibble(Pathway_Name = NA, KEGG_ID = NA, Name = NA, Comment = NA, Time = 'T3')
for(i in 1:length(T3.NMR.diff_table$`KEGG Compound ID`)){
  k_id <- T3.NMR.diff_table$`KEGG Compound ID`[i]
  query <- keggGet(k_id)
  pathway <- query[[1]]$PATHWAY
  if(is.null(pathway)){
    pathway <- NA
  }
  t_tibble <- tibble(Pathway_Name = pathway,
                     KEGG_ID = k_id,
                     Name = T3.NMR.diff_table$Name[i],
                     Comment = T3.NMR.diff_table$Comment[i],
                     Time = 'T3')
  T3_NMR.kegg <- rbind(T3_NMR.kegg, t_tibble)
  print(i)
}
```

Plotting the Heatmap of the KEGG Pathways associated with NMR data

```{r fig.height=10}

T1_NMR_counts.kegg <- T1_NMR.kegg %>% 
  group_by(Comment) %>% 
  count(Pathway_Name) %>% 
  ungroup() %>% 
  filter(n > 2) %>%
  mutate(Time = 'T1')

T2_NMR_counts.kegg <- T2_NMR.kegg %>% 
  group_by(Comment) %>% 
  count(Pathway_Name) %>% 
  ungroup() %>% 
  filter(n > 2) %>%
  mutate(Time = 'T2')

T3_NMR_counts.kegg <- T3_NMR.kegg %>% 
  group_by(Comment) %>% 
  count(Pathway_Name) %>% 
  ungroup() %>% 
  filter(n > 2) %>%
  mutate(Time = 'T3')

nmr_kegg <- rbind(T1_NMR_counts.kegg, T2_NMR_counts.kegg, T3_NMR_counts.kegg) %>% 
  mutate(Source = paste(Time, Comment, sep = '_')) %>% 
  select(-Comment, -Time)

nmr_kegg <- pivot_wider(nmr_kegg, names_from = Source, values_from = n)

nmr_kegg[is.na(nmr_kegg)] <- 0

nmr_kegg <- pivot_longer(nmr_kegg, cols = 2:11, names_to = 'Source', values_to = 'n')

nmr_kegg <- ddply(nmr_kegg, .(Source), transform, rescale = rescale(n))

kegg_nmr_heatmap <- ggplot(nmr_kegg,
                       aes(x = Source,
                           y = Pathway_Name,
                           fill = rescale)) +
  geom_tile(color = 'white') +
  coord_fixed() +
  scale_fill_distiller(palette = 'Blues', direction = 1) +
  theme_bw() +
  theme(plot.title = element_text(face = 'bold', hjust = 0.5),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        axis.title.x = element_blank()) +
  labs(title = 'Dysregulated KEGG Pathways',
       y = 'Pathway Name',
       fill = 'Scaled counts')
kegg_nmr_heatmap

figure_file <- file.path(figures_dir, 'KEGG_tile_heatmap_with_NMR.png')
ggsave(figure_file, kegg_nmr_heatmap, height = 10)
```

## 3.4 KEGG Pathways from the sum of LC-MS2 and NMR data

```{r fig.height=10}
# Join both datasets together
T1_all_counts <- T1.kegg %>% 
  select(Pathway_Name, KEGG_ID, Name, Comment, Time) %>% 
  rbind(T1_NMR.kegg) %>%
  group_by(Comment) %>% 
  count(Pathway_Name) %>% 
  ungroup() %>% 
  filter(n > 1) %>%
  mutate(Time = 'T1')

T2_all_counts <- T2.kegg %>% 
  select(Pathway_Name, KEGG_ID, Name, Comment, Time) %>% 
  rbind(T2_NMR.kegg) %>%
  group_by(Comment) %>% 
  count(Pathway_Name) %>% 
  ungroup() %>% 
  filter(n > 1) %>%
  mutate(Time = 'T2')

T3_all_counts <- T3.kegg %>% 
  select(Pathway_Name, KEGG_ID, Name, Comment, Time) %>% 
  rbind(T3_NMR.kegg) %>%
  group_by(Comment) %>% 
  count(Pathway_Name) %>% 
  ungroup() %>% 
  filter(n > 1) %>%
  mutate(Time = 'T3')

all_kegg <- rbind(T1_all_counts, T2_all_counts, T3_all_counts) %>% 
  mutate(Source = paste(Time, Comment, sep = '_')) %>% 
  select(-Comment, -Time)

all_kegg <- pivot_wider(all_kegg, names_from = Source, values_from = n)

all_kegg[is.na(all_kegg)] <- 0

all_kegg <- pivot_longer(all_kegg, cols = 2:12, names_to = 'Source', values_to = 'n')

all_kegg <- ddply(all_kegg, .(Source), transform, rescale = rescale(n))

kegg_all_heatmap <- ggplot(all_kegg,
                       aes(x = Source,
                           y = Pathway_Name,
                           fill = rescale)) +
  geom_tile(color = 'white') +
  coord_fixed() +
  scale_fill_distiller(palette = 'Blues', direction = 1) +
  theme_bw() +
  theme(plot.title = element_text(face = 'bold', hjust = 0.5),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        axis.title.x = element_blank()) +
  labs(title = 'Dysregulated KEGG Pathways',
       y = 'Pathway Name',
       fill = 'Scaled counts')
kegg_all_heatmap

figure_file <- file.path(figures_dir, 'KEGG_tile_heatmap_all_data.png')
ggsave(figure_file, kegg_all_heatmap, height = 10)
```

## 3.5 Extract vectors with KEGG IDs

```{r}
lc_kegg_ids <- rbind(T1.kegg, T2.kegg, T3.kegg) %>% 
  select(KEGG_ID, Time, Comment) %>% 
  mutate(Source = paste(Time, Comment, sep = '_')) %>% 
  select(-Comment, - Time) %>% 
  separate_rows(KEGG_ID, sep = '; ') %>% 
  group_by(Source) %>% 
  distinct() %>% 
  ungroup()
  
nmr_kegg_ids <- rbind(T1_NMR.kegg, T2_NMR.kegg, T3_NMR.kegg) %>% 
  filter(!is.na(KEGG_ID)) %>% 
  select(KEGG_ID, Time, Comment) %>% 
  mutate(Source = paste(Time, Comment, sep = '_')) %>% 
  select(-Comment, - Time) %>% 
  group_by(Source) %>% 
  distinct() %>% 
  ungroup()

all_kegg_ids <- rbind(lc_kegg_ids, nmr_kegg_ids) %>% 
  mutate(Color = case_when(Source == 'T1_Not present in control' ~ '#FF3300',
                           Source == 'T1_Upregulated' ~ '#FF6600',
                           Source == 'T1_Downregulated' ~ '#FF9900',
                           Source == 'T1_Only present in control' ~ '#FFCC00',
                           Source == 'T2_Not present in control' ~ '#000066',
                           Source == 'T2_Upregulated' ~ '#000099',
                           Source == 'T2_Downregulated' ~ '#3333CC',
                           Source == 'T2_Only present in control' ~ '#6969FF',
                           Source == 'T3_Not present in control' ~ '#005C2E',
                           Source == 'T3_Upregulated' ~ '#3329A3c',
                           Source == 'T3_Downregulated' ~ '#66CC99',
                           Source == 'T3_Only present in control' ~ '#99FFCC')) %>% 
  mutate(for_kegg_mapper = paste(KEGG_ID, Color, sep = ' '))

table_file <- file.path(tables_dir, 'KEGG_ids_for_mapper.csv')
write_csv(all_kegg_ids, table_file)
```


# 4. Analyze Metabolika Pathways at each time point

```{r}
T1.metabolika <- T1.diff_table %>% 
  select(`Metabolika Pathways`, pval.adj, Comment) %>% 
  filter(pval.adj < 0.05) %>% 
  separate_rows(`Metabolika Pathways`, sep = ';') %>% 
  filter(!is.na(`Metabolika Pathways`)) %>% 
  group_by(Comment) %>% 
  count(`Metabolika Pathways`) %>% 
  ungroup() %>% 
  mutate(Time = 'T1')

T2.metabolika <- T2.diff_table %>% 
  select(`Metabolika Pathways`, pval.adj, Comment) %>% 
  filter(pval.adj < 0.05) %>% 
  separate_rows(`Metabolika Pathways`, sep = ';') %>% 
  filter(!is.na(`Metabolika Pathways`)) %>% 
  group_by(Comment) %>% 
  count(`Metabolika Pathways`) %>% 
  ungroup() %>% 
  mutate(Time = 'T2')

T3.metabolika <- T3.diff_table %>% 
  select(`Metabolika Pathways`, pval.adj, Comment) %>% 
  filter(pval.adj < 0.05) %>% 
  separate_rows(`Metabolika Pathways`, sep = ';') %>% 
  filter(!is.na(`Metabolika Pathways`)) %>% 
  group_by(Comment) %>% 
  count(`Metabolika Pathways`) %>% 
  ungroup() %>% 
  mutate(Time = 'T3')
```

```{r}
all_metabolika <- rbind(T1.metabolika, T2.metabolika, T3.metabolika) %>% 
  mutate(Source = paste(Time, Comment, sep = '_')) %>% 
  select(-Comment, -Time)

all_metabolika <- pivot_wider(all_metabolika, names_from = Source, values_from = n)

all_metabolika[is.na(all_metabolika)] <- 0

all_metabolika <- pivot_longer(all_metabolika, cols = 2:3, names_to = 'Source', values_to = 'n')

all_metabolika <- ddply(all_metabolika, .(Source), transform, rescale = rescale(n))

metabolika_all_heatmap <- ggplot(all_metabolika,
                       aes(x = Source,
                           y = Metabolika.Pathways,
                           fill = rescale)) +
  geom_tile(color = 'white') +
  coord_fixed() +
  scale_fill_distiller(palette = 'Blues', direction = 1) +
  theme_bw() +
  theme(plot.title = element_text(face = 'bold', hjust = 0.5),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        axis.title.x = element_blank()) +
  labs(title = 'Dysregulated metabolika Pathways',
       y = 'Pathway Name',
       fill = 'Scaled counts')
metabolika_all_heatmap

figure_file <- file.path(figures_dir, 'metabolika_tile_heatmap_all_data.png')
ggsave(figure_file, metabolika_all_heatmap)
```



