---
title: "Differential Analysis"
author: "Author"
date: "3/19/2021"
output: html_document
---

This Notebook is to perform the differential analysis of the normalized AUC from *Compound Discoverer*.

# 1. Importing Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggrepel)
library(ggsci)
library(pheatmap)
library(RColorBrewer)
library(readxl)
library(UpSetR)
library(ggnewscale)
library(ggpubr)
library(DiagrammeR)
source('functions_cdis_diff.R')
source('functions_cdis_exploration.R')
```

# 2. Import data

Because differential analysis includes the calculations of means, the data to be used in this section is the normalized untransformed data

```{r set_path, message=FALSE, warning=FALSE}
# set path variables
project_dir <- getwd()
project_name <- 'project'
figures_dir <- file.path(project_dir, paste0(project_name, '_output_figures'))
tables_dir <- file.path(project_dir,  paste0(project_name, '_output_tables'))
norm_auc_table_file <- file.path(tables_dir, 'normalized_transformed_auc_table.csv')

# If the flag was setted before, nothing needs to be changed here, the correct file will be automatically used
if(label == TRUE){
  compounds_table_file <- file.path(tables_dir, 'compounds_table.csv')
}else{
  compounds_table_file <- file.path(tables_dir, 'gap_filled_compounds_table.csv')
}

# Load auc_table

norm.matrix <- read_csv(norm_auc_table_file) %>% 
  column_to_rownames(var = '...1')

# Load compounds table and add a column that has names for identified compounds and FeatureID for the rest (for plotting purposes)

compounds_table <- read_csv(compounds_table_file)

# Import metadata and fix names
metadata_file <- file.path(tables_dir, 'fixed_metadata.csv')
metadata <- read_csv(metadata_file)

```

# 3. Calculate differentially expressed metabolites at different time points


## 3.1 Obtaining sample lists for all comparisons

```{r}

# Get the average AUC per each of the treatments
## Get samples per treatment

t0.samples <- metadata %>% filter(treatment == 'CTR', time == 'T0') %>% pull(SampleID)
t1.ctr.samples <- metadata %>% filter(treatment == 'CTR', time == 'T1') %>% pull(SampleID)
t2.ctr.samples <- metadata %>% filter(treatment == 'CTR', time == 'T2') %>% pull(SampleID)
t3.ctr.samples <- metadata %>% filter(treatment == 'CTR', time == 'T3') %>% pull(SampleID)
t4.ctr.samples <- metadata %>% filter(treatment == 'CTR', time == 'T4') %>% pull(SampleID)

t1.wp.samples <- metadata %>% filter(treatment == 'WP', time == 'T1') %>% pull(SampleID)
t2.wp.samples <- metadata %>% filter(treatment == 'WP', time == 'T2') %>% pull(SampleID)
t3.wp.samples <- metadata %>% filter(treatment == 'WP', time == 'T3') %>% pull(SampleID)
t4.wp.samples <- metadata %>% filter(treatment == 'WP', time == 'T4') %>% pull(SampleID)

```

## 3.2 DE at 24 hours after the water pulse

### 3.2.1 Calculate differential expression table

```{r}

de.t1.c_wp <- get_diff_table(norm.matrix, 
                             control.sample_list = t1.ctr.samples, 
                             treatment.sample_list = t1.wp.samples, 
                             log2_transformed = TRUE) %>% 
  left_join(select(compounds_table, -SampleID, -AUC), by = 'FeatureID') %>% 
  distinct() %>% 
  mutate(Comment = case_when(log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

table_file <- file.path(tables_dir, 'de.t1.c_wp.csv')
write_csv(de.t1.c_wp, table_file)
```


### 3.2.2 Volcano plot

```{r}

lfc.t <- 1
pval.t <- 0.05

volcano.t1.c_wp <- plot_volcano(de.t1.c_wp, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'DE metabolites at T1 - Wp vs Control') +
  scale_y_continuous(limits = c(0, 4)) +
  scale_x_continuous(limits = c(-4, 4))
volcano.t1.c_wp

figure_file <- file.path(figures_dir, 'volcano.t1.c_wp.png')
ggsave(figure_file, volcano.t1.c_wp, dpi = 300)

```

## 3.3 DE at T2 after the water pulse

### 3.3.1 Calculate differential expression table

```{r}

de.t2.c_wp <- get_diff_table(norm.matrix, 
                             control.sample_list = t2.ctr.samples, 
                             treatment.sample_list = t2.wp.samples, 
                             log2_transformed = TRUE) %>% 
  left_join(select(compounds_table, -SampleID, -AUC), by = 'FeatureID') %>% 
  distinct() %>% 
  mutate(Comment = case_when(log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

table_file <- file.path(tables_dir, 'de.t2.c_wp.csv')
write_csv(de.t2.c_wp, table_file)
```


### 3.3.2 Volcano plot

```{r}

lfc.t <- 1
pval.t <- 0.05

volcano.t2.c_wp <- plot_volcano(de.t2.c_wp, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'DE metabolites at T2 - Wp vs Control') +
  scale_y_continuous(limits = c(0, 4)) +
  scale_x_continuous(limits = c(-4, 4))
volcano.t2.c_wp

figure_file <- file.path(figures_dir, 'volcano.t2.c_wp.png')
ggsave(figure_file, volcano.t2.c_wp, dpi = 300)

```

## 3.4 DE at T3 after the water pulse

### 3.4.1 Calculate differential expression table

```{r}

de.t3.c_wp <- get_diff_table(norm.matrix, 
                             control.sample_list = t3.ctr.samples, 
                             treatment.sample_list = t3.wp.samples, 
                             log2_transformed = TRUE) %>% 
  left_join(select(compounds_table, -SampleID, -AUC), by = 'FeatureID') %>% 
  distinct() %>% 
  mutate(Comment = case_when(log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

table_file <- file.path(tables_dir, 'de.t3.c_wp.csv')
write_csv(de.t3.c_wp, table_file)
```


### 3.4.2 Volcano plot

```{r}

lfc.t <- 1
pval.t <- 0.05

volcano.t3.c_wp <- plot_volcano(de.t3.c_wp, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'DE metabolites at t3 - Wp vs Control') +
  scale_y_continuous(limits = c(0, 4)) +
  scale_x_continuous(limits = c(-4, 4))
volcano.t3.c_wp

figure_file <- file.path(figures_dir, 'volcano.t3.c_wp.png')
ggsave(figure_file, volcano.t3.c_wp, dpi = 300)

```

## 3.5 DE at T4 after the water pulse

### 3.5.1 Calculate differential expression table

```{r}

de.t4.c_wp <- get_diff_table(norm.matrix, 
                             control.sample_list = t4.ctr.samples, 
                             treatment.sample_list = t4.wp.samples, 
                             log2_transformed = TRUE) %>% 
  left_join(select(compounds_table, -SampleID, -AUC), by = 'FeatureID') %>% 
  distinct() %>% 
  mutate(Comment = case_when(log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

table_file <- file.path(tables_dir, 'de.t4.c_wp.csv')
write_csv(de.t4.c_wp, table_file)
```

### 3.5.2 Volcano plot

```{r}

lfc.t <- 1
pval.t <- 0.05

volcano.t4.c_wp <- plot_volcano(de.t4.c_wp, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'DE metabolites at t4 - Wp vs Control') +
  scale_y_continuous(limits = c(0, 4)) +
  scale_x_continuous(limits = c(-4, 4))
volcano.t4.c_wp

figure_file <- file.path(figures_dir, 'volcano.t4.c_wp.png')
ggsave(figure_file, volcano.t4.c_wp, dpi = 300)

```

# 4 Calculate differentially expressed metabolites between timepoints

## 4.1 Control - T1 vs T0

### 4.1.1 Calculate differential expression table

```{r}

de.ctr.t0_t1 <- get_diff_table(norm.matrix, 
                             control.sample_list = t0.samples, 
                             treatment.sample_list = t1.ctr.samples, 
                             log2_transformed = TRUE) %>% 
  left_join(select(compounds_table, -SampleID, -AUC), by = 'FeatureID') %>% 
  distinct() %>% 
  mutate(Comment = case_when(log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

table_file <- file.path(tables_dir, 'de.ctr.t0_t1.csv')
write_csv(de.ctr.t0_t1, table_file)
```

### 4.1.2 Volcano plot

```{r}

lfc.t <- 1
pval.t <- 0.05

volcano.ctr.t0_t1 <- plot_volcano(de.ctr.t0_t1, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'DE metabolites CTR - T1 vs T0') +
  scale_y_continuous(limits = c(0, 4)) +
  scale_x_continuous(limits = c(-4, 4))
volcano.ctr.t0_t1

figure_file <- file.path(figures_dir, 'volcano.ctr.t0_t1.png')
ggsave(figure_file, volcano.ctr.t0_t1, dpi = 300)

```

## 4.2 Control - T2 vs T0

### 4.2.1 Calculate differential expression table

```{r}

de.ctr.t0_t2 <- get_diff_table(norm.matrix, 
                             control.sample_list = t0.samples, 
                             treatment.sample_list = t2.ctr.samples, 
                             log2_transformed = TRUE) %>% 
  left_join(select(compounds_table, -SampleID, -AUC), by = 'FeatureID') %>% 
  distinct() %>% 
  mutate(Comment = case_when(log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

table_file <- file.path(tables_dir, 'de.ctr.t0_t2.csv')
write_csv(de.ctr.t0_t2, table_file)
```

### 4.2.2 Volcano plot

```{r}

lfc.t <- 1
pval.t <- 0.05

volcano.ctr.t0_t2 <- plot_volcano(de.ctr.t0_t2, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'DE metabolites CTR - t2 vs T0') +
  scale_y_continuous(limits = c(0, 4)) +
  scale_x_continuous(limits = c(-4, 4))
volcano.ctr.t0_t2

figure_file <- file.path(figures_dir, 'volcano.ctr.t0_t2.png')
ggsave(figure_file, volcano.ctr.t0_t2, dpi = 300)

```

## 4.3 Control - T3 vs T0

### 4.3.1 Calculate differential expression table

```{r}

de.ctr.t0_t3 <- get_diff_table(norm.matrix, 
                             control.sample_list = t0.samples, 
                             treatment.sample_list = t3.ctr.samples, 
                             log2_transformed = TRUE) %>% 
  left_join(select(compounds_table, -SampleID, -AUC), by = 'FeatureID') %>% 
  distinct() %>% 
  mutate(Comment = case_when(log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

table_file <- file.path(tables_dir, 'de.ctr.t0_t3.csv')
write_csv(de.ctr.t0_t3, table_file)
```

### 4.3.2 Volcano plot

```{r}

lfc.t <- 1
pval.t <- 0.05

volcano.ctr.t0_t3 <- plot_volcano(de.ctr.t0_t3, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'DE metabolites CTR - t3 vs T0') +
  scale_y_continuous(limits = c(0, 4)) +
  scale_x_continuous(limits = c(-4, 4))
volcano.ctr.t0_t3

figure_file <- file.path(figures_dir, 'volcano.ctr.t0_t3.png')
ggsave(figure_file, volcano.ctr.t0_t3, dpi = 300)

```


## 4.4 Control - T4 vs T0

### 4.4.1 Calculate differential expression table

```{r}

de.ctr.t0_t4 <- get_diff_table(norm.matrix, 
                             control.sample_list = t0.samples, 
                             treatment.sample_list = t4.ctr.samples, 
                             log2_transformed = TRUE) %>% 
  left_join(select(compounds_table, -SampleID, -AUC), by = 'FeatureID') %>% 
  distinct() %>% 
  mutate(Comment = case_when(log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

table_file <- file.path(tables_dir, 'de.ctr.t0_t4.csv')
write_csv(de.ctr.t0_t4, table_file)
```

### 4.4.2 Volcano plot

```{r}

lfc.t <- 1
pval.t <- 0.05

volcano.ctr.t0_t4 <- plot_volcano(de.ctr.t0_t4, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'DE metabolites CTR - t4 vs T0') +
  scale_y_continuous(limits = c(0, 4)) +
  scale_x_continuous(limits = c(-4, 4))
volcano.ctr.t0_t4

figure_file <- file.path(figures_dir, 'volcano.ctr.t0_t4.png')
ggsave(figure_file, volcano.ctr.t0_t4, dpi = 300)

```

## 4.5 WP - T1 vs T0

### 4.5.1 Calculate differential expression table

```{r}

de.wp.t0_t1 <- get_diff_table(norm.matrix, 
                              control.sample_list = t0.samples, 
                              treatment.sample_list = t1.wp.samples, 
                              log2_transformed = TRUE) %>% 
  left_join(select(compounds_table, -SampleID, -AUC), by = 'FeatureID') %>% 
  distinct() %>% 
  mutate(Comment = case_when(log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

table_file <- file.path(tables_dir, 'de.wp.t0_t1.csv')
write_csv(de.wp.t0_t1, table_file)
```

### 4.5.2 Volcano plot

```{r}

lfc.t <- 1
pval.t <- 0.05

volcano.wp.t0_t1 <- plot_volcano(de.wp.t0_t1, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'DE metabolites WP - t1 vs T0') +
  scale_y_continuous(limits = c(0, 4)) +
  scale_x_continuous(limits = c(-4, 4))
volcano.wp.t0_t1

figure_file <- file.path(figures_dir, 'volcano.wp.t0_t1.png')
ggsave(figure_file, volcano.wp.t0_t1, dpi = 300)

```

## 4.6 WP - T2 vs T0

### 4.6.1 Calculate differential expression table

```{r}

de.wp.t0_t2 <- get_diff_table(norm.matrix, 
                              control.sample_list = t0.samples, 
                              treatment.sample_list = t2.wp.samples, 
                              log2_transformed = TRUE) %>% 
  left_join(select(compounds_table, -SampleID, -AUC), by = 'FeatureID') %>% 
  distinct() %>% 
  mutate(Comment = case_when(log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

table_file <- file.path(tables_dir, 'de.wp.t0_t2.csv')
write_csv(de.wp.t0_t2, table_file)
```

### 4.6.2 Volcano plot

```{r}

lfc.t <- 1
pval.t <- 0.05

volcano.wp.t0_t2 <- plot_volcano(de.wp.t0_t2, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'DE metabolites WP - t2 vs T0') +
  scale_y_continuous(limits = c(0, 4)) +
  scale_x_continuous(limits = c(-4, 4))
volcano.wp.t0_t2

figure_file <- file.path(figures_dir, 'volcano.wp.t0_t2.png')
ggsave(figure_file, volcano.wp.t0_t2, dpi = 300)

```

## 4.7 WP - T3 vs T0

### 4.7.1 Calculate differential expression table

```{r}

de.wp.t0_t3 <- get_diff_table(norm.matrix, 
                              control.sample_list = t0.samples, 
                              treatment.sample_list = t3.wp.samples, 
                              log2_transformed = TRUE) %>% 
  left_join(select(compounds_table, -SampleID, -AUC), by = 'FeatureID') %>% 
  distinct() %>% 
  mutate(Comment = case_when(log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

table_file <- file.path(tables_dir, 'de.wp.t0_t3.csv')
write_csv(de.wp.t0_t3, table_file)
```

### 4.7.2 Volcano plot

```{r}

lfc.t <- 1
pval.t <- 0.05

volcano.wp.t0_t3 <- plot_volcano(de.wp.t0_t3, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'DE metabolites WP - t3 vs T0') +
  scale_y_continuous(limits = c(0, 4)) +
  scale_x_continuous(limits = c(-4, 4))
volcano.wp.t0_t3

figure_file <- file.path(figures_dir, 'volcano.wp.t0_t3.png')
ggsave(figure_file, volcano.wp.t0_t3, dpi = 300)

```

## 4.8 WP - T4 vs T0

### 4.8.1 Calculate differential expression table

```{r}

de.wp.t0_t4 <- get_diff_table(norm.matrix, 
                              control.sample_list = t0.samples, 
                              treatment.sample_list = t4.wp.samples, 
                              log2_transformed = TRUE) %>% 
  left_join(select(compounds_table, -SampleID, -AUC), by = 'FeatureID') %>% 
  distinct() %>% 
  mutate(Comment = case_when(log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

table_file <- file.path(tables_dir, 'de.wp.t0_t4.csv')
write_csv(de.wp.t0_t4, table_file)
```

### 4.8.2 Volcano plot

```{r}

lfc.t <- 1
pval.t <- 0.05

volcano.wp.t0_t4 <- plot_volcano(de.wp.t0_t4, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'DE metabolites WP - t4 vs T0') +
  scale_y_continuous(limits = c(0, 4)) +
  scale_x_continuous(limits = c(-4, 4))
volcano.wp.t0_t4

figure_file <- file.path(figures_dir, 'volcano.wp.t0_t4.png')
ggsave(figure_file, volcano.wp.t0_t4, dpi = 300)

```

### Volcano mixed

```{r}
volcano_treatment <- ggarrange(volcano.t1.c_wp, volcano.t2.c_wp, volcano.t3.c_wp, volcano.t4.c_wp, ncol = 4)
volcano_treatment <- annotate_figure(volcano_treatment, top = text_grob('Comparisons of treatments per time', 
                                                                        face = 'bold', size = 20))
volcano_ctr <- ggarrange(volcano.ctr.t0_t1, volcano.ctr.t0_t2, volcano.ctr.t0_t3, volcano.ctr.t0_t4, ncol = 4)
volcano_ctr <- annotate_figure(volcano_ctr, top = text_grob('Comparisons of different times within control samples', 
                                                                        face = 'bold', size = 20))
volcano_wp <- ggarrange(volcano.wp.t0_t1, volcano.wp.t0_t2, volcano.wp.t0_t3, volcano.wp.t0_t4, ncol = 4)
volcano_wp <- annotate_figure(volcano_wp, top = text_grob('Comparisons of different times within WP samples', 
                                                                        face = 'bold', size = 20))

volcano_all <- ggarrange(volcano_treatment, volcano_ctr, volcano_wp, nrow = 3)

figure_file <- file.path(figures_dir, 'volcano.all.png')
ggsave(figure_file, volcano_all, dpi = 300, width = 18, height = 10)

```


## Summary graph

`to be added`

# 5. Hierarchical clustering analysis using heatmaps

Multiple heatmaps will be plotted, based on the differential expression and if the compounds were or not assigned and structure

### Normalized Area under the curve (AUC) heatmap for all detected features

```{r warning=FALSE}

# Initialize graphical device
dev.off()

# Set sampleID as row.names to annotate heatmap

col_annot <- metadata %>% 
  filter(!is.na(treatment)) %>% 
  select(-`Sample Type`) %>% 
  column_to_rownames(var = 'SampleID')
mapcolor <- colorRampPalette(brewer.pal(11, 'RdYlBu'))(100)[100:1]

figure_file <- file.path(figures_dir, 'All_features_heatmap.pdf')

annot_colors <- list(
  time = c(T0 = "#374E55FF", T1 = "#DF8F44FF", T2 = "#00A1D5FF", T3 = "#B24745FF", T4 = "#79AF97FF"),
  treatment = c(CTR ="#1B9E77", WP ="#7570B3"))

pdf(figure_file)
pheatmap(norm.matrix,
         clustering_distance_rows = 'correlation',
         clustering_distance_cols = 'correlation',
         scale = 'row',
         annotation_col = col_annot,
         annotation_colors = annot_colors,
         color = mapcolor,
         show_rownames = FALSE,
         cutree_cols = 2,
         main = 'All features (normalized AUC)'
)
dev.off()
```

[Scaled Area under the curve (AUC) heatmap for all detected features](`r toString(figure_file)`)


