---
title: "RP Lab Incubations - Data Exploration"
author: "Christian Ayala & Viviana Freire"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

# 1. Importing Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(ggpubr)
library(ggsci)
library(UpSetR)
library(rstatix)
source('functions_cdis_exploration.R')
```

# 2. Import data

```{r set_path, message=FALSE}
#set path variables
project_dir <- getwd()

# Give a name for your project so the results are all grouped together in a directory with the same name
project_name <- 'RP_lab_3-4-22'
figures_dir <- file.path(project_dir, paste0(project_name, '_output_figures'))
tables_dir <- file.path(project_dir,  paste0(project_name, '_output_tables'))

# Create output directories
dir.create(figures_dir, showWarnings = FALSE)

dir.create(tables_dir, showWarnings = FALSE)
```

Set if the data to be used is going to be labeled or unlabeled

```{r}
# Flag for labeled / unlabeled data, set TRUE or FALSE

label = FALSE
```


The data for in this script is the **Compounds table** table exported from **Compound Discoverer**. For this script to work export and excel table with the columns: **Name**, **Formula**, **Annot. deltaMass [ppm]**, **Calc. MW**, **RT [min]**, **MS2**, **Area**, **Gap Status** and **Gap Fill Status**.
The **metadata table** can be exported directly from the **Samples tab** in **Compound Discoverer** by copying to a blank text file.


```{r import_data, message=FALSE}
# Import data tables
cd_results_file <- file.path(project_dir, '..', '..', 'data', 'RP_lab_join_3-4-22.xlsx')
cd_results_table <- read_xlsx(cd_results_file)

# Select columns needed for downstream analysis

cd_results_table <- cd_results_table %>%
  filter(!is.na(Formula)) %>% 
  filter(abs(`Annot. DeltaMass [ppm]`) < 5) %>% 
  arrange(desc(`Calc. MW`)) %>% 
  mutate(FeatureID = paste0('Feature',formatC(n():0001, width = 4, flag = '0'))) %>%
  select(FeatureID, Name, Formula, `Calc. MW`, contains('Annotation source'), contains('Results'), contains('Pathways'), 
         contains('Area:'), contains('Gap Status:'), contains('Labeling Status:'), contains('Gap Fill Status:')) %>% 
  # Differentiate between features that share the same name using "peak#" at the end of the name
  group_by(Name) %>% 
  add_count(Name) %>% 
  # Create variable with names for plotting (useful in following scripts)
  mutate(name4plot = ifelse(is.na(Name), FeatureID, ifelse(n == 1, Name, paste0(Name, '-peak', n():1)))) %>% 
  select(-n) %>% 
  ungroup()

# Import metadata and fix names
metadata_file <- file.path(project_dir, '..', '..', 'data', 'metadata_RP_lab.csv')
metadata <- read_csv(metadata_file)

# Select only the useful columns and fix column names

metadata <- metadata %>%  
  rename(SampleID = `Sample Identifier`)

table_file <- file.path(tables_dir, 'fixed_metadata.csv')
write_csv(metadata, table_file)
```

Get the label status of each feature (labeled analysis only)

```{r}
if(label == TRUE){
  label_status <- cd_results_table %>% 
    select(FeatureID, contains('Labeling Status:'))
  
  colnames(label_status) <- str_remove(colnames(label_status), 'Saleska_')
  colnames(label_status) <- str_remove(colnames(label_status), '_23.*')
  
  label_status[label_status == 'Contaminating mass' | label_status == 'Low pattern fit' | label_status == 'Not detected'] <- 'Unlabeled'
  label_status[label_status == 'No warnings' | label_status == 'Irregular exchange' ] <- 'Labeled'
  
  table_file <- file.path(tables_dir, 'label_status.csv')
  write_csv(label_status, table_file)
}
```



# 3. Data Manipulation and thermodynamic indices calculations

```{r data, message=FALSE, warning=FALSE}

# Split formula column into elemental counts
cd_results_table <- separate_formula(cd_results_table)

# Calculate ratios and thermodynamic indices
cd_results_table <- calc_ratios_n_idxs(cd_results_table)

# Calculate classes
cd_results_table <- calc_classes(cd_results_table)

# Gather area under the curve (AUC) values per sample

compounds_table <- cd_results_table %>% 
  select(-contains('Gap Status:'), -contains('Labeling Status:'), -contains('Gap Fill Status:')) %>% 
  pivot_longer(contains('Area:'), names_to = 'SampleID', values_to = 'AUC') %>% 
  filter(AUC > 0) %>%  
  mutate(SampleID = str_remove(SampleID, 'Area: '),
         SampleID = str_remove(SampleID, '.raw.*'))

# Save gap-filled table to be used in Statistical Analysis

table_file <- file.path(tables_dir, 'gap_filled_compounds_table.csv')
write_csv(compounds_table, table_file)

```

### Gap filtering (only for unlabel data)

```{r}

# Gather "Gap Status" for filtering 
if(label == FALSE){
  gap_status <- cd_results_table %>% 
    select(FeatureID, contains('Gap Status:')) %>% 
    gather(contains('Gap Status:'), key = 'SampleID', value = 'gap_status')
  gap_status$SampleID <- str_remove(gap_status$SampleID, 'Gap Status: ')
  gap_status$SampleID <- str_remove(gap_status$SampleID, '.raw.*')
  
  ## Filtering
  
  compounds_table <- left_join(compounds_table, gap_status, by = c('FeatureID', 'SampleID')) %>% 
    filter(gap_status != 'Full gap')
  
  # Plotting types of gaps detected
  
  gap_fill_status <- cd_results_table%>% 
    select(FeatureID, Name, contains('Gap Fill Status:')) %>% 
    pivot_longer(contains('Gap Fill Status:'), names_to = 'SampleID', values_to = 'gap_fill') %>% 
    mutate(SampleID = str_remove(SampleID, 'Gap Fill Status: '),
           SampleID = str_remove(SampleID, '.raw.*')) %>% 
    mutate(gap_fill_type = case_when(gap_fill == 32 ~ 'Filled by spectrum noise',
                                     gap_fill == 128 ~ 'Filled by re-detected peak',
                                     gap_fill == 0 ~ 'No gap to fill',
                                     gap_fill == 16 ~ 'Filled by simulated peak',
                                     gap_fill == 8 ~ 'Filled by trace area',
                                     gap_fill == 64 ~ 'Filled by matching ion')) %>% 
    mutate(gap_fill_type = factor(gap_fill_type, levels = c('No gap to fill', 'Filled by re-detected peak', 
                                                            'Filled by simulated peak', 'Filled by matching ion',
                                                            'Filled by trace area', 'Filled by spectrum noise')))
}

```


``` {r}
# Add metadata information

compounds_table <- left_join(compounds_table, metadata, by = 'SampleID')

table_file <- file.path(tables_dir, 'compounds_table.csv')
write_csv(compounds_table, table_file)

```

# 4. Summary plots

## 4.1. Number of identified compounds

This figure shows the number of compounds that were detected in each file, and how many of them were assigned a structure through database searches

```{r n_identified_masses, message=FALSE}
# Create a new tibble to count the number of compounds per file
summary_table <- compounds_table %>% 
  filter(!is.na(treatment)) %>% 
  select(SampleID, Name) %>% 
  mutate(Status = ifelse(is.na(Name), 'No identified', 'Identified')) %>% 
  group_by(SampleID) %>% 
  count(Status) %>% 
  left_join(metadata, by = c('SampleID')) %>% 
  mutate(Status = factor(Status, levels = c('No identified', 'Identified'))) %>% 
  filter(!is.na(treatment))

id_summary_plot <- plot_col(summary_table, SampleID, n, Status) +
  labs(title = 'Number of masses found',
       x = 'Sample name',
       y = 'Count') +
  theme(plot.title = element_text(face = 'bold', hjust = 0.5),
        axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  facet_grid(cols = vars(treatment), scales = 'free_x', space = 'free_x')

id_summary_plot

figure_file <- file.path(figures_dir, 'id_summary.png')
ggsave(figure_file, id_summary_plot, dpi = 300)

```

## 4.3 Gap filled compounds

```{r}

gap_summary_table <- gap_fill_status %>% 
  select(SampleID, gap_fill_type) %>% 
  group_by(SampleID) %>% 
  count(gap_fill_type) %>% 
  left_join(metadata, by = 'SampleID') %>% 
  filter(!is.na(treatment))

gap_summary_plot <- plot_col(gap_summary_table, SampleID, n , gap_fill_type) +
  labs(title = 'Gap filled compounds',
       x = 'Sample name',
       y = 'Count') +
  theme(plot.title = element_text(face = 'bold', hjust = 0.5),
        axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  scale_fill_manual(values = c('green4', 'aquamarine3', 'blue2', 'blue', 'purple', 'firebrick2')) +
  facet_grid(cols = vars(treatment), scales = 'free_x', space = 'free_x')

gap_summary_plot

figure_file <- file.path(figures_dir, 'gap_summary.png')
ggsave(figure_file, gap_summary_plot, dpi = 300)

```


## 4.2. Upset plots of number of detected compounds

### Compounds detected in each treatment (without gap filling)

```{r message=FALSE}

# Venn diagram of the compounds present in each time
my_list <- list(Control = unique(compounds_table$FeatureID[compounds_table$treatment == 'CTR']),
                Water_pulse = unique(compounds_table$FeatureID[compounds_table$treatment == 'WP']))

figure_file <- file.path(figures_dir, 'upset_per_treatment.png')
png(figure_file, res = 300, width = 2400, height = 1600)
upset(fromList(my_list), order.by = 'freq')
dev.off()
```

### Compounds detected at each time point (without gap filling)

```{r message=FALSE}

# Venn diagram of the compounds present in each time
my_list <- list(T0 = unique(compounds_table$FeatureID[compounds_table$time == 'T0']),
                T1 = unique(compounds_table$FeatureID[compounds_table$time == 'T1']),
                T2 = unique(compounds_table$FeatureID[compounds_table$time == 'T2']),
                T3 = unique(compounds_table$FeatureID[compounds_table$time == 'T3']),
                T4 = unique(compounds_table$FeatureID[compounds_table$time == 'T4']))

figure_file <- file.path(figures_dir, 'upset_per_time.png')
png(figure_file, res = 300, width = 2400, height = 1600)
upset(fromList(my_list), order.by = 'freq')
dev.off()
```

# 5. Exploratory plots

## 5.1. Van Krevelen Diagrams

### Van Krevelen diagram by treatment


```{r message=FALSE}

# Ven Krevelen Diagram based on time
vank <- compounds_table %>% 
  filter(!is.na(treatment)) %>% 
  plot_vank(., Class, treatment)
vank

figure_file <- file.path(figures_dir, 'Venk_diagram_treatment.jpg')
ggsave(figure_file, vank, dpi = 300)

```

### Van Krevelen diagram by time

```{r message=FALSE}

# Ven Krevelen Diagram based on time
vank <- compounds_table %>% 
  filter(!is.na(treatment)) %>% 
  plot_vank(., Class) +
  facet_wrap(~ time)
vank

figure_file <- file.path(figures_dir, 'Venk_diagram_time.jpg')
ggsave(figure_file, vank, dpi = 300)

```

## 6.1 Gibbs Free Energy (GFE) exploratory plots

### Boxplot of GFE of different treatments

```{r message=FALSE}

stat_table <- compounds_table %>% 
  filter(!is.na(treatment)) %>%
  select(GFE, treatment) %>% 
  wilcox_test(GFE ~ treatment) %>% 
  add_xy_position(step.increase = 3) %>% 
  add_significance()

gfe_boxplot <- compounds_table %>% 
  filter(!is.na(treatment)) %>%
  ggplot(aes(x = treatment,
             y = GFE,
             fill = treatment)) +
  geom_boxplot() +
  scale_fill_brewer(palette = 'Dark2') +
  stat_pvalue_manual(data = stat_table, label = 'p.signif', inherit.aes = FALSE, hide.ns = TRUE) +
  labs(title = 'Gibbs Free Energy (GFE) per habitat') +
  theme_bw() +
  theme(plot.title = element_text(face = 'bold', hjust = 0.5))

gfe_boxplot

figure_file <- file.path(figures_dir, 'gfe_treatment.png')
ggsave(figure_file, gfe_boxplot, dpi = 300)
```

### Boxplot of GFE at different times per treatment

```{r message=FALSE}

stat_table <- compounds_table %>% 
  filter(!is.na(treatment)) %>%
  group_by(treatment) %>% 
  wilcox_test(GFE ~ time) %>% 
  add_xy_position(step.increase = 3) %>% 
  add_significance()

gfe_boxplot <- compounds_table %>% 
  filter(!is.na(treatment)) %>%
  ggplot(aes(x = time,
             y = GFE,
             fill = time)) +
  geom_boxplot() +
  scale_fill_jama() +
  stat_pvalue_manual(data = stat_table, label = 'p.adj.signif', inherit.aes = FALSE, hide.ns = TRUE) +
  labs(title = 'Gibbs Free Energy (GFE) per habitat') +
  theme_bw() +
  theme(plot.title = element_text(face = 'bold', hjust = 0.5)) +
  facet_grid(rows = vars(treatment))

gfe_boxplot

figure_file <- file.path(figures_dir, 'gfe_time_per_treatment.png')
ggsave(figure_file, gfe_boxplot, dpi = 300)
```

### Boxplot of GFE per treatment at different times

```{r message=FALSE}

stat_table <- compounds_table %>% 
  filter(!is.na(treatment),
         time != 'T0') %>%
  group_by(time) %>% 
  wilcox_test(GFE ~ treatment) %>% 
  add_xy_position(step.increase = 3) %>% 
  adjust_pvalue() %>% 
  add_significance()

gfe_boxplot <- compounds_table %>% 
  filter(!is.na(treatment),
         time != 'T0') %>%
  ggplot(aes(x = treatment,
             y = GFE,
             fill = treatment)) +
  geom_boxplot() +
  scale_fill_brewer(palette = 'Dark2') +
  stat_pvalue_manual(data = stat_table, label = 'p.adj.signif', inherit.aes = FALSE, hide.ns = TRUE) +
  labs(title = 'Gibbs Free Energy (GFE) per habitat') +
  theme_bw() +
  theme(plot.title = element_text(face = 'bold', hjust = 0.5)) +
  facet_grid(cols = vars(time))

gfe_boxplot

figure_file <- file.path(figures_dir, 'gfe_time_per_treatment.png')
ggsave(figure_file, gfe_boxplot, dpi = 300)
```

